---
title: "text2sdg: Detecting UN Sustainable Development Goals in Text"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{text2sdg}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height = 4
)
library(text2sdg)
```

## Introduction

The United Nations’ Sustainable Development Goals (SDGs) have become an important guideline for organizations to monitor and plan their contributions to social, economic, and environmental transformations. Existing approaches to identify efforts addressing each of the 17 goals rely on economic indicators or, in the case of academic output, on search engines of academic databases. `text2sdg` is the first open source, multi-system tool to detect SDGs in text.     

The `text2sdg` package consists of five functions: `detect_sdg()`, `detect_sdg_systems()`, `detect_any()`, `plot_sdg()`, and `crosstab_sdg()`. The function `detect_sdg()` carries out the detection of SDGs in text using a trained machine learning model, whereas `detect_sdg_systems()` does so using up to six different query systems. The function `detect_any()` enables search for custom search queries. Finally, the functions `plot_sdg()` and `crosstab_sdg()` help visualize and analyze the resulting SDG matches.

## Detecting SDGs using `detect_sdg()`

The `detect_sdg()` function identifies SDGs in texts that are provided via the `text` argument. Inputs to `text` must either be a character vector or an object of `tCorpus` from package `corpustools`. `text` is the only non-default argument of the function, which means that the function can be run with only the texts as input. The output of the function is a `tibble` with one row per match including the following columns (and types):

- `document` (`factor`) - index of element in the character vector or corpus supply for `text`
- `sdg` (`character`) - labels indicating the matched SDGs
- `system` (`character`) - the query system that produced the match
- `hit` (`numeric`) - running index of matches

The `detect_sdg()` function implements a ensemble model based on a random forest architecture that pools the predictions of six labeling systems generated using `detect_sdg_systems()` and also considers text length. The ensemble model can be run for any subset of the 17 SDGs by using the `sdgs` argument. Our research shows that the ensemble model implemented by `detect_sdg()` outperforms any individual labeling system and the OSDG tool developed by the UN SDG AI Lab (Wulff, Meier, & Mata, 2023).   

The example below runs the `detect_sdg()` function for the `projects` data set included in the package and prints the results. The data set is a character vector containing 500 descriptions of randomly selected University of Basel research projects that were funded by the Swiss National Science Foundation (<a href="https://p3.snf.ch">https://p3.snf.ch</a>). The analysis produced a total of `62` matches.   

```{r, eval=FALSE}
# detecting SDGs in projects
hits_default <- detect_sdg(projects)
hits_default
```


## Detecting SDGs using `detect_sdg_systems()`

The `detect_sdg_systems()` function generates SDG labels by using up to six different SDG labeling systems. The function works similarly to `detect_sdg()` but has additional arguments to control which systems are run and the aggregation of the output. The output is a `tibble` with one row per match including the following columns (and types):

- `document` (`factor`) - index of element in the character vector or corpus supply for `text`
- `sdg` (`character`) - labels indicating the matched SDGs
- `system` (`character`) - the query system that produced the match
- `query_id` (`integer`) - identifier of query in the query system
- `features` (`character`) - words in the document that were matched by the query
- `hit` (`numeric`) - running index of matches for each query system

The example below runs the `detect_sdg_systems()` function for the `projects` data. The analysis produced a total of `835` matches using the four default query systems Aurora, Elsevier, Auckland, and SIRIS.   

```{r, eval=FALSE}
# detecting SDGs in projects
hits_default <- detect_sdg_systems(projects)
hits_default
```

### Selecting query systems 

By default `sdg_detect_systems()` runs the four query systems, Aurora, Elsevier, Auckland, and SIRIS. Using the function's `system` argument the user can control which systems are run. There are two additional query systems that can be selected, SDSN and SDGO. These two systems are much simpler and less restrictive than the former four as they only rely on basic keyword matching, whereas the four default systems are based on more complex search queries. As a result, the four default systems are more specific but also more prone to misses, whereas the two optional systems are more sensitive but also more prone to false alarms. Our research suggests that the four default systems are likely more accurate in balanced datasets including SDG-related and SDG-unrelated content (Wulff, Meier, & Mata, 2023). More information about the systems can be gathered from the help files of the respective query data frames, `aurora_queries`, `elsevier_queries`, `auckland_queries`, `siris_queries`, `sdsn_queries`, and `sdgo_queries`.

The example below runs the `detect_sdg_systems()` on the `projects` using all query systems, including the two keyword-based systems. The resulting `tibble`s reveal that Aurora is most conservative (60 hits), followed by SIRIS (166), Elsevier (235), Auckland (374), and then by large margin the two keyword-based systems SDSN (2,589) and SDGO (3,629). Note that the high numbers for the two keyword-based systems imply that, on average, 5 and 7 SDGs, respectively, are identified per document. 

```{r, echo = F}
options(tibble.print_min = 5)
```


```{r}
# detecting SDGs using all query systems
hits_all <- detect_sdg_systems(projects,
  system = c("Aurora", "Elsevier", "Auckland", "SIRIS", "SDSN", "SDGO")
)

# count hits of systems
table(hits_all$system)
```

```{r, echo = F}
options(tibble.print_min = 10)
```

### Selecting SDGs

By default the `detect_sdg()` and `detect_sdg_systems()` will try to detect all 17 SDGs. However, the user can limit the set of SDGs sought in text using the `sdgs` argument, which takes a numeric vector with integers in [1,17] as input. When using the `detect_sdg_systems()` function, the user should note that Elsevier, SIRIS and Auckland contain queries only for the first 16 SDGs, exluding queries for goal 17 - Global Partnerships for the Goals.

The example below runs the `detect_sdg_systems()` function only for SDGs 1, 2, 3, 4, and 5. 

```{r}
# detecting only for SDGs 1 to 5
hits_sdg_subset <- detect_sdg_systems(projects, sdgs = 1:5)
hits_sdg_subset
```

### Controlling the output

By default, `detect_sdg_systems()` returns matches at the level of query. If a system has multiple queries for a single SDG, the output can include multiple hits (and rows) per document and system. Separating hits by queries can be useful because different queries typically capture different aspects of a given SDGs, which will be revealed through the keywords that were matched by the queries. These keyword matches are shown in the column `features` and, hence, we refer to thsi type of output as `"features"`. If the user is not interested in separating matches by queries and only cares about matches at the level of documents, a reduced output can be selected by setting the `output` argument to `"documents"`. In this case, the `detect_sdg_systems()` returns a `tibble` that includes only distinct matches of `document`, `system`, and `sdg` combinations, concatenates the values of `features` into a single character string, and drops the column `query_id`.  

The example below shows the alternative output resulting from setting `output = "documents"`. 

```{r}
# return documents output format
detect_sdg_systems(projects, output = "documents")
```

### Keeping track of progress

By default the `detect_sdg()` and `detect_sdg_systems()` function prints messages to communicate the progress of the underlying SDG detection process, which can take several minutes. To suppress these messages, the user can set the `verbose` argument to `FALSE`. 

## Custom search with `detect_any()`

The `detect_any()` function permits specification of custom query systems. The function operates similarly to `detect_sdg_systems()`, but it requires an additional argument `queries` that specifies the queries to be employed. The `queries` argument expects a `tibble` with the following columns:

- `system` (`character`) - names used to label query systems.
- `queries` (`character`) - queries used in search. 
- `sdg` (`integer`) - mapping of queries to SDGs. 

The queries in the custom query set can be Lucene-style queries following the syntax of the `corpustools` package. See `vignette("corpustools")`. This is illustrated in the example below. First, a tibble of three queries is defined that includes a single system and three queries that are mapped onto two sdgs, 3 and 7. The first query represents a simple keyword search, whereas queries 2 and 3 are proper search queries using logical operators.  


```{r}
# definition of query set
my_queries <- tibble::tibble(
  system = "my_system",
  query = c("theory", "analysis OR analyses OR analyzed", "study AND hypothesis"),
  sdg = c(3, 7, 7)
)

# return documents output format
detect_any(
  text = projects,
  system = my_queries
)
```


## Visualizing hits with `plot_sdg()`

To visualize the hits produced by the `detect_*` functions, the `text2sdg` package provides the function `plot_sdg()`. The function produces barplots illustrating either the hit frequencies produced by the different query systems. It is built on the `ggplot2` package, which provides high levels of flexibility for adapting and extending its visualizations.  

By default `plot_sdg()` produces a stacked barplot of absolute hit frequencies. Frequencies are determined on the document level for each system and SDGs present in the `tibble` that was provided to the function's `hits` argument. If multiple hits per document, system and SDG combination exist, the function returns a message of how many duplicate hits have been suppressed.  

The example below produces the default visualization for the hits of all five systems. Since the object was created with `"output = features"`, the function informs that a total of 2490 duplicate hits were removed. 

```{r}
# show stacked barplot of hits
plot_sdg(hits_all)
```

### Adjusting visualizations

The `plot_sdg()` function has several arguments that permit adjustment of the visualization. The `systems` and `sdgs` arguments can be used to visualize subsets of systems and/or SDGs. The `normalize` argument can be used to normalize the absolute frequency  by the number of documents (`normalize = "documents"`) or by the total number of hits within a system (`normalize = "systems"`). The `color` argument can be used to adapt the color set used for the systems. The `sdg_titles` argument can be used to add the full titles of the SDGs. The `remove_duplicates` argument can be used to retain any duplicate hits of document, system, and SDG combinations. Finally, the `...` arguments can be used to pass on additional arguments to the `geom_bar()` function that underlies `detect_sdg_systems()`.      

The example below uses some of the available arguments to make adjustments to the default visualization. With `normalize = "systems"` and `position = "dodge"`, an argument passed to `geom_bar()`, it shows the proportion of SDG hits per system with bars presented side-by-side rather than stacked. Furthermore, due to `sdg_titles = TRUE` the full titles are shown rather than SDG numbers.  

```{r}
# show normalized, side-by-side barplot of hits
plot_sdg(hits_all,
  sdg_titles = TRUE,
  normalize = "systems",
  position = "dodge"
)
```

### Extending visualizations with `ggplot2`

Because `plot_sdg()` is implemented using `ggplot2`, visualizations can easily be extended using functions from the `ggplot2` universe. The example below illustrates this. Using the `facet_wrap` function separate panels are created, one for each system, that show the absolute frequencies of hits per SDG. 

```{r}
# show system hits in separate panels
plot_sdg(hits_all) +
  ggplot2::facet_wrap(~system, ncol = 1, scales = "free_y")
```


## Analyzing hits using `crosstab_sdg()`

To assist the user in understanding the relationships among SDGs and query systems, the `text2sdg` package provides the `crosstab_sdg()` function. The function takes as input the `tibble` of hits produced by any of the detect functions and compares hits between either systems or SDGs. Comparing hits by system means that correlations are determined across all documents and all SDGs for every pair of systems to produce a fully crossed table of system correlations. Conversely, comparing hits by SDG means that correlations are determined across all documents and all systems for every pair of SDGs to produce a fully crossed table of SDG correlations. 

### Correspondence between query systems

By default the `crosstab_sdg()` function compares `systems`, which is illustrated below for the hits for all five systems. Note that the `crosstab_sdg()` function only considers distinct combinations of documents, systems, and SDGs implying that the output type of `detect_sdg_systems()` does not matter; it will automatically treat the hits as if they had been produced using `output = documents`. 

The analysis reveals two noteworthy results. First, correlations between systems are overall rather small. Second, query systems are more similar to systems of the same type, i.e., query or keyword-based.  

```{r}
# evaluate correspondence between systems
crosstab_sdg(hits_all) %>% round(2)
```

When `crosstab_sdg()` evaluates the correspondence between query systems it does not distinguish between hits of different SDGs. Correlations for individual SDGs could be different from the overall correlations, and it is likely that they are higher, on average. To determine the correspondence between query systems for individual SDGs, the user can use the `sdgs` argument. For instance, `sdgs = 1` will only result in a comparison of systems using only hits of SDG 1. 

### Correspondence between SDGs

The `crosstab_sdg()` can also be used to analyze, in a similar fashion, the correspondence between SDGs. To do this, the `compare` argument must be set to `"sdgs"`. Again, correlations are calculated for distinct hits, while ignoring, in this case, the systems from which the hits originated. 

The example below analyzes the correspondence of all SDGs across all systems. The resulting cross table reveals strong correspondences between certain pairs of SDGs, such as, for instance, between `SDG-01` and `SDG-02` or between `SDG-07` and `SDG-13`. 

```{r}
# evaluate correspondence between systems
crosstab_sdg(hits_all, compare = "sdgs") %>% round(2)
```

