---
title: "text2sdg: An open-source solution to monitoring sustainable development goals from text"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{text2sdg}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, fig.height=4
)
library(text2sdg)

```

## Introduction

The United Nations’ Sustainable Development Goals (SDGs) have become an important guideline for organizations to monitor and plan their contributions to social, economic, and environmental transformations. Existing approaches to identify efforts addressing each of the 17 goals rely on economic indicators or, in the case of academic output, on search engines of academic data bases. `text2sdg` is the first open source, multi-system tool to detect SDGs in text.     

The `text2sdg` package consists of four functions: `detect_sdg()`, `detect_any()`, `plot_sdg()`, and `crosstab_sdg()`. The function `detect_sdg()` carries out the detection of SDGs in text using up to five different established search systems. The function `detect_any()` enables search for custom search queries. Finally, the functions `plot_sdg()` and `crosstab_sdg()` help visualize and analyze the resulting SDG matches.

## Detecting SDGs using `detect_sdg()`

The `detect_sdg()` identifies SDGs in texts that are provided via the `text` argument. Inputs to `text` must either be a character vector or an object of `tCorpus` from package `corpustools`. `text` is the only non-default argument of the function, which means that the function can be run with only the texts as input. The output of the function is a `tibble` with one row per match including the following columns (and types):

- `document` (`factor`) - index of element in the character vector or corpus supplie for `text`
- `sdg` (`character`) - labels indicating the matched SDGs
- `system` (`character`) - the search system that produced the match
- `query_id` (`integer`) - identifier of query in the search system
- `features` (`character`) - words in the document that were matched by the query
- `hit` (`numeric`) - running index of matches for each search system

The example below runs the `detect_sdg()` for the `projects` data set included in the package prints the results. The data set is character vector containing 500 descriptions of University of Basel research projects that were funded by the Swiss National Science Foundation (<a href="https://p3.snf.ch">https://p3.snf.ch</a>). The analysis produced a total of `462` matches using the three default searchs systems Aurora, Elsevier, and SIRIS.   

```{r}
# detecting SDGs in projects
hits_default <- detect_sdg(projects)
hits_default
```

### Selecting search systems 

By default `sdg_detect()` runs the three query-based search systems, Aurora, Elsevier, and SIRIS. Using the function's `system` argument the user can control which systems are run. There are two additional keyword-based systems that can be selected, Ontology and SDSN. In comparison, the two keyword-based systems tend to identify larger numbers of SDGs compared to the query-based search systems. Systematic validations are outstanding, however, it is likely that keyword-based systems produce a larger number of false positives, which is the reason they are not included in the function's default search system set. 

The example below runs the `detect_sdg()` on the `projects` for each search system individually, including the two keyword baed system. The resulting `tibble`s reveal that Aurora is most conservative (60 hits), followed by SIRIS (167), Elsevier (235), and then with large margin the two keyword-based systems SDSN (2,575) and Ontology (3,617). Note that the high numbers for the two keyword-based systems imply that, on average, 5 respectively 7 SDGs are identified per document. 

```{r, echo = F}
options(tibble.print_min = 5)
```

```{r}
# detecting SDGs using all available search systems
hits_all <- detect_sdg(projects, 
                       system = c("aurora", "elsevier", "siris", "sdsn", "ontology"))

# count hits of systems
table(hits_all$system)
```

```{r, echo = F}
options(tibble.print_min = 10)
```

### Selecting SDGs

By default the `detect_sdg()` function aims to detect all SDGs covered by the different systems, which is all 17 for Aurora, Ontology, and SDSN, and 16 for Elsevier and Siris. The latter two do not include queries for goal 17 - Global Partnerships for the Goals. If the user is only interested in a subset of SDGs or wants to reduce runtime, the `sdgs` argument can be used to run subsets of SDGs. The argument takes a numeric vector with integers in [1,17] indiciating the SDGs that shall be detected. 

The example below runs the `detect_sdg()` function only for SDGs 1, 2, 3, 4, and 5. 

```{r}
# detecting only for SDGs 1 to 5
hits_sdg_subset <- detect_sdg(projects, sdgs = 1:5)
hits_sdg_subset
```

### Controlling the ouptut

By default the output of the `detect_sdg()` function returns matches at the level of query. This means one SDG can be matched multiple times in one document which is possible because there multiple queries belonging to the same SDGs produce in all systems but Elsevier. On this level, it is meaningful to study the features (individual words) that were matched by the individual queries, which is why we refer to this output type as `"features"`. If the user is only interested in matches at the level of documents, a reduced output can be selected by setting the `output` argument to `"documents"`. As a result, the `detect_sdg()` returns a `tibble` that includes only distinct matches of `document`, `system`, and `sdg` combinations and drops the columns `query_id` and `features`.  

The example below shows the alternative output resulting from setting `output = "documents"`. 

```{r}
# return documents output format
detect_sdg(projects, output = "documents")
```

### Keeping track of progress

By default the `detect_sdg()` function prints messages whenever it begins running a search system. This can help track the progress of the function, which for large number of texts can take several minutes. To suppress these messages, the user can set the `verbose` argument to `FALSE`. 

## Custom search with `detect_any()`

The `text2sdg` provides with `detect_any()` a function that permits specification of custom search systems. The function operates similarly to `detect_sdg()`, but it requires an additional argument `queries` that specifies the queries to be employed. The `queries` argument expects a `tibble` with the following columns:

- `system` (`character`) - names used to label search systems.
- `queries` (`character`) - queries used in search. 
- `sdg` (`integer`) - mapping of queries to SDGs. 

The queries in the custom query set can be Lucene-style queries following the syntax of the `corpustools` package. See `vignette("corpustools")`. This is illustrated in the example below. First, a tibble of three queries is defined that includes a single system and three queries that are mapped onto two sdgs, 3 and 7. The first query represents a simple keyword search, whereas queries 2 and 3 are proper search queries using logical operators.  


```{r}
# definition of query set
my_queries <- tibble::tibble(system = "my_system",
                              query = c("theory", "analysis OR analyses OR analyzed", "study AND hypothesis"),
                              sdg = c(3,7,7))

# return documents output format
detect_any(text = projects, 
           queries = my_queries)
```


## Visualizing hits with `plot_sdg()`

To visualize the hits produced by `detect_sdg`, the `text2sdg` package provides the function `plot_sdg()`. The function produces barplots illustrating either the hit frequencies produced by the different search systems. It is built on the `ggplot2` package, which provides high levels of flexibility for adapting and extending its visualizations.  

By default `plot_sdg()` produces a stacked barplot of absolute hit frequencies. Frequencies are determined on the document level for each system and sdgs present in the `tibble` that was provided to the function's `hits` argument. If multiple hits per document, system and SDG combination exist, the function returns a message of how many duplicate hits have been suppressed.  

The example below produces the default visualization for the hits of all five systems. Since the object was created with `"output = features"`, the function informs that a total of 2490 duplicate hits were removed. 

```{r}
# show stacked barplot of hits
plot_sdg(hits_all)
```

### Adjusting visualizations

The `plot_sdg()` has several arguments that permit adjustment of the visualization. The `systems` and `sdgs` arguments can be used to visualize subsets of systems and/or SDGs. The `normalize` argument can be used to normalize the absolute frequency  by the number of documents (`normalize = "documents"`) or by the total number of hits within a system (`normalize = "systems"`). The `color` argument can be used to adapt the color set used for the systems. The `sdg_titles` argument can be used to add the full titles of the SDGs. The `remove duplicates` argument can be used to retain any duplicates hits of document, system, and SDG combinations. Finally, the `...` arguments can be used to pass on additional arguments to the `geom_bar()` function that underlies `plot_sdg()`.      

The example below uses some of the available arguments to make adjustments to the default visualization. With `normalize = "systems"` and `position = "dodge"`, an argument passed to `geom_bar()`, it shows the proportion of SDG hits per system with bars presented side-by-side rather than stacked. Furthermore, due to `sdg_titles = TRUE` the full titles are shown rather than SDG numbers.  

```{r}
# show normalized, side-by-side barplot of hits
plot_sdg(hits_all, 
         sdg_titles = TRUE,
         normalize = "systems", 
         position = "dodge")
```

### Extending visualizations with `ggplot2`

Since `plot_sdg()` is implemented using `ggplot2`, visualizations can easily be extended using functions from the `ggplot2` universe. The example below illustrates this. Using the `facet_wrap` function separate panels are created, one for each system, that show the absolute frequencies of hits per SDG. 

```{r}
# show system hits in separate panels
plot_sdg(hits_all) + 
  ggplot2::facet_wrap(~system, ncol= 1, scales = "free_y")
```


## Analyzing hits using `crosstab_sdg()`

To assist the user in understanding the relationships among SDGs and search systems, the `text2sdg` package provides the `crosstab_sdg()` function. The function takes as input the `tibble` of hits produced by `detect_sdg()` and compares hits either between systems or SDGs. Comparing hits by systems means that correlations are determined across all documents and all SDGs for every pair of systems to produce a fully crossed table of system correlations. Conversely, comparing hits by SDG means that correlations are determined across all documents and all systems for every pair of SDGs to produce a fullly crossed table of SDG correlations. 

### Correspondence between search systems

By default the `crosstab_sdg()` function compares `systems`, which is illustrated below for the hits for all five systems. Note that the `crosstab_sdg()` function only considers distinct combinations of documents, systems, and SDGs implying that the output type of `detect_sdg()` does not matter; it will automatically treat the hits as if they had been produced using `output = documents`. 

The analysis reveals two noteworthy results. First, correlations between systems are overall rather small. Second, search systems are more similar to systems of the same type, i.e., query or keyword-based.  

```{r}
# evaluate correspondence between systems 
crosstab_sdg(hits_all) %>% round(2)
```

When `crosstab_sdg()` evaluates the correspondence between search systems it does not distinguish between hits of different SDGs. Correlations for individual SDGs could be different from the overall correlations, and it is likely that they are higher, on average. To determine the correspondence between search systems for individual SDGs, the user can use the `sdgs` argument. For instance, `sdgs = 1` will only result in a comparisons of systems using only hits of SDG 1. 

### Correspondence between SDGs

The `crosstab_sdg()` can also be used to analyze, in a similar fashion, the correspondence between SDGs. To do this, the `compare` argument must be set to `"sdgs"`. Again, correlations are calculated for distinct hits, while ignoring, in this case, the systems from which the hits originated. 

The example below analyzes the correspondence of all SDGs across all systems. The resulting cross table reveals strong correspondences between certain pairs of SDGs, such as, for instance, between `SDG-01` and `SDG-02` or between `SDG-07` and `SDG-13`. 

```{r}
# evaluate correspondence between systems 
crosstab_sdg(hits_all, compare = "sdgs") %>% round(2)

```

